---
title: "Prognostic Predictive Signature"
author: "Pranjal Vaidya"
date: "`r Sys.Date()`"
linkcolor: blue
output:
  rmdformats::readthedown:
    highlight: kate
    number_sections: true
    code_folding: show
---

# Introduction 

Generating Radiomic Signature based on the LASSO GLMNET model for predicting Disease-Free Survival in Early-Stage NSCLC. 

# Data Load/Merge

Initial Setup and Package Loads in R 

Packages used for the analysis.
```{r initial_setup, cache=FALSE, message = FALSE, warning = FALSE}
library(glmnet);library(survival);library(survminer);library(readxl);library(ggplot2); library(GGally);library(knitr); library(rmdformats); library(magrittr);
library(skimr); library(Hmisc); library(Epi); library(vcd); library(tidyverse) 

## Global options

options(max.print="75")
opts_chunk$set(comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)


skimr::skim_with(numeric = list(hist = NULL),
                 integer = list(hist = NULL))
```

## Loading the Raw Data into R 

Loading raw dataset into R.

```{r load_train_set}
train <- read.csv("sample.csv")
train %>%
  select(time) %>%
  summary()
```

## Feature Selection 

LASSO Feature Selection Method.
alpha=0 gives ridge regression and alpha =1 is LASSO regression.

input to the cross validation LASSO model
x1: Feature Matrix (Rows: patients and col: features)
y1: time (months/days/years), status_vector(1:censor, 2:dead)

output: 
10 fold cross-validation LASSO plot 
```{r Feature Selection}
x1 <- train[,(8:length(train))]
x <- data.matrix(x1, rownames.force = NA)
y1 <- train[,(2:3)]
y <- data.matrix(y1, rownames.force = NA)


library(survival)
cvfit = cv.glmnet(x, y,family = 'cox',alpha=0.1)
plot(cvfit)
```

Seelcting top features.
output:
coeffs_arranges: matrix of name of features and corresponding coefficients
names: top feature names
my_coefficients: coefficients of top features

```{r top_features}
tmp_coeffs <- coef(cvfit, s = "lambda.min")
non_zero_coefs <- data.frame(name = tmp_coeffs@Dimnames[[1]][tmp_coeffs@i + 1], coefficient = (tmp_coeffs@x))
coeffs_arranges <- non_zero_coefs[order(-non_zero_coefs$coefficient),]
names <- coeffs_arranges[,1]
my_coefficients <- (coeffs_arranges[,2])
my_coefficients
```


# Contruction of QuRiS

Multivariate Analysis with LASSO top features on training cohort. 
```{r Mulivariate_all_features}
varnames = sapply(1:length(names), function(i){
  (paste0(names[i]))
})

iformula <- as.formula(sprintf("Surv(time, status) ~ %s ", paste(varnames, collapse='+')))  
res.cox <- coxph(iformula, data = train)
summary(res.cox)
```

Creating a signature using LASSO coefficients: 

```{r creating_signature}
train_set <- lapply(1:length(varnames), function(i) {
  ifor <-  my_coefficients[i] 
  k <- (sprintf("%s", varnames[i]))
  feature_list <- train[,k]
  value11 <- feature_list*ifor
  df <- data.frame(value11)
})


store <- data.frame(train_set)
QuRiS <- rowSums(store)
```


# Entire dataset

## Multivariate Model
Multivariate analysis with the signature and calculating CI with signature alone:

```{r signature}
train <- cbind(train, QuRiS)

res.cox <- coxph(Surv(time, status) ~ QuRiS, data = train)
summary(res.cox)
res.cox[["coefficients"]]
```

Threshold was selected usign the MATLAB function in the repository. For example here, we used just the median value of the constructed QuRiS 
```{r dividing_data_based_median}
threshold <- median(train$QuRiS)
train$group <- 'Group1'
train$group[train$QuRiS >= threshold] <- 'Group2'
threshold
```

## Kaplan_Meier Survival Curve
Kaplan-Meier plot based on predicted high and low risk groups
```{r KM plot for training}
fit2 <- survfit(Surv(time, status) ~ group, data = train)

ggsurvplot(
   fit2,                     # survfit object with calculated statistics.
   data = train,             # data used to fit survival curves.
   size = 1.2,
   palette = c("#FFCC33", "#0080FF"),
   risk.table = TRUE,       # show risk table.
   pval = TRUE,             # show p-value of log-rank test..
   xlab = "Time in months" ,  # customize X axis label.
   cumcensor = TRUE      # plot the number of censored subjects at time t
 #  ncensor.plot.height = 0.25
)

```

## Hazard Ratio
```{r hazard ratio}
variables <- c("group")
formula <- sapply(variables,
                        function(x) as.formula(paste('Surv(time, status)~', x)))

univariate_analysis <- lapply(formula, function(x){coxph(x, data = train)})
# Extract data 
results <- lapply(univariate_analysis,
                       function(x){ 
                         x <- summary(x)
                         p.value<-signif(x$wald["pvalue"], digits=4)
                         beta<-signif(x$coef[1], digits=4);#coeficient beta
                         HR <-signif(x$coef[2], digits=4);#exp(beta)
                         HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                         HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                         HR <- paste0(HR, " (", 
                                      HR.confint.lower, "-", HR.confint.upper, ")")
                         res<-c(beta, HR, p.value)
                         names(res)<-c("beta", "HR (95% CI for HR)", 
                                       "p.value")
                         return(res)
                       })
res <- t(as.data.frame(results, check.names = FALSE))
as.data.frame(res)
```

```{r}
write.csv(train, file = "dataset_train.csv")
```

# Surgery Alone group (therapy = 0)


Taking only cases for which therapy is euqal to 0 (i.e. has not received any additional treatment)

```{r surgery_group}
surgery <- with(train, subset(train, !train$Cases %in% train$Cases[is.na(time) | therapy == 1]))
```

## Multivariate Model
```{r multivariate_surgery}
res.cox <- coxph(Surv(time, status) ~ QuRiS, data = surgery)
summary(res.cox)
res.cox[["coefficients"]]
```

## Kaplan-Meier Curve
Kaplan-Meier plot based on predicted high and low risk groups
```{r KM plot for surgery}
fit2 <- survfit(Surv(time, status) ~ group, data = surgery)

ggsurvplot(
   fit2,                     # survfit object with calculated statistics.
   data = surgery,             # data used to fit survival curves.
   size = 1.2,
   palette = c("#FFCC33", "#0080FF"),
   risk.table = TRUE,       # show risk table.
   pval = TRUE,             # show p-value of log-rank test..
   xlab = "Time in months" ,  # customize X axis label.
   cumcensor = TRUE      # plot the number of censored subjects at time t
 #  ncensor.plot.height = 0.25
)

```

## Hazard Ratio
```{r hazard ratio surgery}
variables <- c("group")
formula <- sapply(variables,
                        function(x) as.formula(paste('Surv(time, status)~', x)))

univariate_analysis <- lapply(formula, function(x){coxph(x, data = surgery)})
# Extract data 
results <- lapply(univariate_analysis,
                       function(x){ 
                         x <- summary(x)
                         p.value<-signif(x$wald["pvalue"], digits=4)
                         beta<-signif(x$coef[1], digits=4);#coeficient beta
                         HR <-signif(x$coef[2], digits=4);#exp(beta)
                         HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                         HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                         HR <- paste0(HR, " (", 
                                      HR.confint.lower, "-", HR.confint.upper, ")")
                         res<-c(beta, HR, p.value)
                         names(res)<-c("beta", "HR (95% CI for HR)", 
                                       "p.value")
                         return(res)
                       })
res <- t(as.data.frame(results, check.names = FALSE))
as.data.frame(res)
```


# Adjuvant-Chemotherapy Group (therapy = 1)

Taking only cases for whcih therapy is euqal to 1 (i.e. They received any additional treatment)

```{r adj_chemo_group}
adj_chemo <- with(train, subset(train, !train$Cases %in% train$Cases[is.na(time) | therapy == 0]))
```

## Multivariate Model
```{r train_signature}
res.cox <- coxph(Surv(time, status) ~ QuRiS, data = adj_chemo)
summary(res.cox)
res.cox[["coefficients"]]
```

## Kaplan-Meier Curve
Kaplan-Meier plot based on predicted high and low risk groups
```{r KM plot for adj_chemo}
fit2 <- survfit(Surv(time, status) ~ group, data = adj_chemo)

ggsurvplot(
   fit2,                     # survfit object with calculated statistics.
   data = adj_chemo,             # data used to fit survival curves.
   size = 1.2,
   palette = c("#FFCC33", "#0080FF"),
   risk.table = TRUE,       # show risk table.
   pval = TRUE,             # show p-value of log-rank test..
   xlab = "Time in months" ,  # customize X axis label.
   cumcensor = TRUE      # plot the number of censored subjects at time t
 #  ncensor.plot.height = 0.25
)

```

## Hazard Ratio

```{r hazard_ratio_adj.chemo}
variables <- c("group")
formula <- sapply(variables,
                        function(x) as.formula(paste('Surv(time, status)~', x)))

univariate_analysis <- lapply(formula, function(x){coxph(x, data = surgery)})
# Extract data 
results <- lapply(univariate_analysis,
                       function(x){ 
                         x <- summary(x)
                         p.value<-signif(x$wald["pvalue"], digits=4)
                         beta<-signif(x$coef[1], digits=4);#coeficient beta
                         HR <-signif(x$coef[2], digits=4);#exp(beta)
                         HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                         HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                         HR <- paste0(HR, " (", 
                                      HR.confint.lower, "-", HR.confint.upper, ")")
                         res<-c(beta, HR, p.value)
                         names(res)<-c("beta", "HR (95% CI for HR)", 
                                       "p.value")
                         return(res)
                       })
res <- t(as.data.frame(results, check.names = FALSE))
as.data.frame(res)
```

